---
title: 'Can R and Shiny make me a better fisherman? Part 4'
subtitle: 'Exploratory analysis on my fishing data (shiny and plotly)'
author: 'Aurélien Callens'
date: '2022-04-12'
slug: R-shiny-fishing-part4
categories:
  - R
tags:
  - tidyverse
  - shiny
  - plotly

share_img: '/img_post/seabass_bay.jpg'
bigimg: [{src: "/img_post/seabass_bay.jpg", desc: "Seabass"}]
codefolding_show: hide
---

In this post, I explore the data I have collected during the last year with the updated version of the application (presented [here](/2021/06/01/r-shiny-fishing-part3/)). This quick exploratory analysis is performed with two packages I really enjoy: *Plotly* and *shiny*.

For reminder, my new application store the data in three csv files. The first one contains variables related to the fishing conditions. The second one contains information about my catches and finally the third one contains information about the characteristics of the lures I used during the session.

## Shiny to explore fishing data by session

I coded a small shiny application that provide a summary of the tide and river flow conditions, the lure changes and catches for each session. Don't hesitate to explore my fishing data! 

<div class="wrap" style="width: 100%; height: 1500px;
    padding: 0; overflow: hidden;">
<iframe style='width: 800px;
    height: 1550px;
    border: 0;
    -ms-transform: scale(0.25);
    -moz-transform: scale(0.25);
    -o-transform: scale(0.25);
    -webkit-transform: scale(0.25);
    transform: scale(0.95);
    -ms-transform-origin: 0 0;
    -moz-transform-origin: 0 0;
    -o-transform-origin: 0 0;
    -webkit-transform-origin: 0 0;
    transform-origin: 0 0;' src="https://aureliencallens.shinyapps.io/Fishing_analysis/"> </iframe>
</div>

## Code of the shiny application 

Here is the code of the plotly graphs in the application:

```{r, eval=FALSE}
library(plotly)
library(tidyverse)


#' For the tide plot
#' 
#' @param dat first dataframe with session characteristics
#' @param n_ses the id (number) of the session
#' @param temporal_range number of hours to display (before and after the session)
#' @return A plotly object
plot_tide_ses <- function(dat, n_ses, temporal_range = 4){

  dat_t <- dat %>% 
    filter(Session == n_ses) %>% 
    mutate(Tide_ts = list(eval(parse(text = Ts_tide))))
  dat_tide <- as.data.frame(dat_t$Tide_ts)
  dat_tide$hour <- as.POSIXct(dat_tide$hour, origin = "1970-01-01")
  dat_tide$Water <- as.numeric(as.character(dat_tide$Water))
  
  plot_ly(data = dat_tide, 
          x = ~ hour, 
          y = ~ Water, 
          mode = 'lines') %>%  
    layout(shapes = list(
      list(type = 'line',
           x0 = as.POSIXct(dat_t$Beg),
           x1 = as.POSIXct(dat_t$Beg),
           y0 = min(dat_tide$Water),
           y1 = max(dat_tide$Water),
           line = list(dash = 'dot', width = 1)),
      list(type = 'line',
           x0 =  as.POSIXct(dat_t$End),
           x1 = as.POSIXct(dat_t$End),
           y0 = min(dat_tide$Water),
           y1 = max(dat_tide$Water),
           line = list(dash = 'dot', width = 1))),
      xaxis = list(range = as.POSIXct(c(as.POSIXct(dat_t$Beg) - 3600*temporal_range ,
                                        as.POSIXct(dat_t$End) + 3600*temporal_range )),
                   title = ""),
      yaxis = list(title = "Tide level"))
}

#' For the river flow plot
#' 
#' @param dat first dataframe with session characteristics
#' @param n_ses the id (number) of the session
#' @param past_days number of previous to display (before the session)
#' @return A plotly object
plot_flow_ses <- function(dat, n_ses, past_days = 4){
  dat_t <- dat %>% 
    filter(Session == n_ses) %>% 
    mutate(Flow_ts = list(eval(parse(text = Ts_flow))))
  
  dat_flow <- as.data.frame(dat_t$Flow_ts)
  dat_flow$Date <- as.POSIXct(dat_flow$Date, origin = "1970-01-01")
  dat_flow$Nive <- as.numeric(as.character(dat_flow$Nive))
  dat_flow$Adour <- as.numeric(as.character(dat_flow$Adour))
  
  
  dat_flow <- dat_flow %>% 
    pivot_longer(cols = c(Nive, Adour), 
                 names_to = "River",
                 values_to = "Flow")
  
  plot_ly(data = dat_flow, 
          x = ~ Date,
          y = ~ Flow, 
          color = ~ River, 
          mode = 'lines') %>%  
    layout(shapes = list(
      list(type='line',
           x0 = as.POSIXct(dat_t$Beg),
           x1 = as.POSIXct(dat_t$Beg),
           y0 = min(dat_flow$Flow),
           y1 = max(dat_flow$Flow),
           line = list(dash = 'dot', width = 1))),
      xaxis = list(range = as.POSIXct(c(as.POSIXct(dat_t$Beg) - 3600*24*past_days,
                                        as.POSIXct(dat_t$End) )),
                   title = ""))
}

#' Function to prepare the dataset for the plot of lure change and catch
#' 
#' @param lure third dataframe with lure changes (hours) and characteristics
#' @param session first dataframe with session characteristics
#' @param ses_n the id (number) of the session
#' @return A dataframe
start_end_fonction <- function(lure, session, ses_n){
  dat_ses <- session %>% 
    filter(Session == ses_n)
  
  dat_lure <- lure %>% 
    filter(n_ses == ses_n)
  
  startdates <- dat_lure$time
  enddates <- c(startdates[-1], dat_ses$End)
  
  data.frame(change = length(startdates):1, 
             start = as.POSIXct(startdates),
             end = as.POSIXct(enddates),
             type = dat_lure$type_lure,
             text = paste(dat_lure$color_lure, dat_lure$length_lure))
}

#' For the plot of lure change and catch
#' 
#' @param lure third dataframe with lure changes (hours) and characteristics
#' @param caught second dataframe with fish caught characteristics
#' @param session first dataframe with session characteristics
#' @param n_ses the id (number) of the session
#' @return A plotly object
lure_change <- function(lure, caught, dat, n_ses){
  
  df <- start_end_fonction(lure, dat, n_ses)
  
  catch <- caught %>% 
    filter(n_ses == n_ses)
  
  dat_t <- dat %>% 
    filter(Session == n_ses) %>% 
    mutate(Tide_ts = list(eval(parse(text = Ts_tide))))
  dat_tide <- as.data.frame(dat_t$Tide_ts)
  dat_tide$hour <- as.POSIXct(dat_tide$hour, origin = "1970-01-01")
  dat_tide$Water <- as.numeric(as.character(dat_tide$Water))
  
  plot_ly() %>% 
    add_segments(data = df,
                 x = ~ start,
                 xend = ~ end,
                 y = ~ change,
                 yend = ~ change,
                 color = ~ type,
                 #text = ~ text,
                 size = I(5),
                 alpha = 0.8) %>%
    add_segments(x = as.POSIXct(catch$time),
                 xend = as.POSIXct(catch$time),
                 y = min(df$change),
                 yend = max(df$change),
                 line = list(color = "red", dash = "dash"),
                 name = 'Fish caught') %>%
    add_trace(data = dat_tide, 
              x = ~ hour, 
              y = ~ Water, 
              mode = 'lines', 
              yaxis = "y2",
              name = "Water level",
              alpha = 0.4,
              hoverinfo = 'skip'
    ) %>% 
    layout(xaxis = list(range = c(df$start[1] - 1000 , df$end[nrow(df)] + 1000),
                        title = ""),
           yaxis = list(title = "", zeroline = FALSE, showline = FALSE,
                        showticklabels = FALSE, showgrid = FALSE ),
           yaxis2 = list(overlaying = "y", side = "right"))
}

```

Here is the code of this simple yet informative application: 

```{r, eval=FALSE}
library(shiny)
library(shinyWidgets)
library(shinydashboard)
library(plotly)
library(tidyverse)
source('plot_functions.R')
dat <- read_csv("session1.csv")
caught <- read_csv("catch1.csv")
lure <- read_csv("lure.csv")

# In order to save the tide and flow time series I parse the data in the dataframe
# The following line is used to transform the parsed text into usable values
dat_t <- dat %>% 
  mutate(Tide_ts = list(eval(parse(text = Ts_tide))),
         Flow_ts = list(eval(parse(text = Ts_flow))))

body <- dashboardBody(fluidPage(
  # Application title
  h1("Exploratory analysis of fishing data",
  align = "center",
  style = "padding: 40px;  text-align: center;  background: #605ca8;  color: white;  font-size: 40px;"),
  br(),
  # Dropdown menu to select the fishing session
  fluidRow(align = "center",
           pickerInput(inputId = 'Ses',
                       label = h3('Select a fishing session:'),
                       choices = unique(dat$Session[-1]),
                       options = list(
                         style = "btn-primary"),
                       choicesOpt = list(
                         style = rep_len("font-size: 75%; line-height: 1.6;", 4)
                       ))),
  br(),
  br(),
  # Key figures of the session
  fluidRow(
    valueBoxOutput("progressD", width = 4),
    valueBoxOutput("progressF", width = 4),
    valueBoxOutput("progressL", width = 4)),
  br(),
  
  br(),
  # Graphs of the tide and river flow of recent days
  fluidRow(
    box(title = "Tidal water level", status = "primary", 
        plotlyOutput("TidePlot"), width = 6),
    box(title = "River flow", status = "primary",
        plotlyOutput("FlowPlot"), width = 6)),
  br(),
  # Graph lure changes during the session + catch
  fluidRow(
    box(title = "Lures tested and fish capture", status = "warning", 
        plotlyOutput("LurePlot"), width=12))
))

ui <- dashboardPage(
  
  dashboardHeader(disable = TRUE),
  
  dashboardSidebar(disable = TRUE),
  
  body
)


# Define server logic required to draw a histogram
server <- function(input, output) {
  
  # Duration
  output$progressD <- renderValueBox({
    Duration = as.integer(difftime(as.POSIXct(dat$End[dat$Session == input$Ses]), as.POSIXct(dat$Beg[dat$Session == input$Ses]), units = 'mins'))
    valueBox(tags$p("Duration", style = "font-size: 80%;"),
             tags$p(paste(Duration, "min"), style = "font-size: 150%; font-weight: bold;"),
             icon = icon("clock"), color = "purple")
  })
  
  # Number of fish
  
  output$progressF <- renderValueBox({
    fish_caught = as.integer(caught %>% filter(n_ses == input$Ses) %>% nrow())
    valueBox(tags$p("Fish caught", style = "font-size: 80%;"), tags$p(fish_caught, style = "font-size: 150%;font-weight: bold;"),
             icon = icon("trophy"), color = "purple")
  })
  
  # Number of lures tried
  
  output$progressL <- renderValueBox({
    Lure = as.integer(lure %>% filter(n_ses == input$Ses) %>% nrow())
    valueBox(tags$p("Lure tried", style = "font-size: 80%;"), tags$p(Lure, style = "font-size: 150%;font-weight: bold;"),
             icon = icon("fish"), color = "purple")
  })
  
  output$TidePlot <- renderPlotly({
    # generate plot depending on session
    plot_tide_ses(dat, input$Ses, 4)
  })
  output$FlowPlot <- renderPlotly({
    # generate plot depending on session
    plot_flow_ses(dat_t, input$Ses, 4)
  })
  output$LurePlot <- renderPlotly({
    # generate plot depending on session
    lure_change(lure, caught, dat, input$Ses)
  })
  
}

# Run the application 
shinyApp(ui = ui, server = server)

```


